---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/joaobtj/hf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/joaobtj/hf/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/joaobtj/hf/actions/workflows/pkgdown.yaml/badge.svg)](https://joaobtj.github.io/hf/)
[![hf status badge](https://joaobtj.r-universe.dev/badges/hf)](https://joaobtj.r-universe.dev)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)

<!-- badges: end -->



# hf: Head Loss Calculations for Pipelines

The **hf** package provides simple, efficient, and mathematically rigorous functions to calculate friction head loss, flow rate, and required pipe diameters using standard hydraulic equations.

It fully supports the universal **Darcy-Weisbach** equation (with Colebrook-White, Swamee-Jain, and other friction factors) and empirical equations (such as **Hazen-Williams** and **Flamant**), making it an ideal tool for agronomic and engineering hydraulic designs.

## 🚀 Interactive Web Application

You can try the package directly in your browser without installing R! The `hf` package includes a serverless Shiny application powered by WebAssembly (`shinylive`), allowing you to perform complex hydraulic calculations on the fly.

**With the interactive app, you can:**

* **Solve for any variable:** Instantly calculate Head Loss, Flow Rate, or Required Pipe Diameter based on your system parameters.
* **Compare Methods:** Switch seamlessly between the universal Darcy-Weisbach equation (testing both Colebrook-White and Swamee-Jain factors) and the empirical Hazen-Williams formula.
* **Visualize System Dynamics:** View automatically generated plots, including **System Curves** (Head Loss vs. Flow Rate) and **Friction Sensitivity** (Head Loss vs. Pipe Diameter) to understand how your pipeline behaves.
* **Zero Installation:** The app runs entirely client-side in your browser with no R backend required.

👉 **[Try the hf Hydraulic Calculator](https://joaobtj.github.io/hf/app/)**

## 📦 Installation

You can install the development version of `hf` from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("joaobtj/hf")
```

Alternatively, you can install it via R-Universe:

```r
install.packages("hf", repos = c("https://joaobtj.r-universe.dev", "https://cloud.r-project.org"))


```

## 💡 Usage

### Darcy-Weisbach Equation

The universal Darcy-Weisbach method allows for the calculation of head loss using different friction factor functions. By default, it uses the implicit Colebrook-White equation.

```{r example-darcy}
library(hf)

# Calculate head loss (m) for a 100m pipe, 0.02 m³/s flow, 0.1m diameter
calc_head_loss_darcy(
  length = 100,
  flow = 0.02,
  diameter = 0.1,
  roughness = 0.00026
)

# Use functional injection to apply the Swamee-Jain approximation
calc_head_loss_darcy(
  length = 100,
  flow = 0.02,
  diameter = 0.1,
  roughness = 0.00026,
  friction_fun = calc_friction_sj
)
```



## 📖 Documentation

For full documentation and bilingual tutorials (English and Portuguese), visit the package website: **[joaobtj.github.io/hf](https://joaobtj.github.io/hf/)**

## 🤝 Citation & Contact

Author: Joao Batista Tolentino Jr. (UFSC - Campus Curitibanos)  
Contact: [joao.tolentino@ufsc.br](mailto:joao.tolentino@ufsc.br) | ORCID: [0000-0003-1303-4202](https://orcid.org/0000-0003-1303-4202)
